The Community Land Model, version 5.1 \\ One-at-a-time Parameter Perturbation Ensemble

### 2023-05-22 ms/ms.tex

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{../figs/variance.pdf}
\caption{needs caption}
\label{fig:violins}
\end{figure}

@@

In this project, we have identified and perturbed 188 CLM parameters to create a large one-at-a-time parameter perturbation ensemble. This ensemble is useful for understanding parametric controls on CLM processes and has already been leveraged as a starting point for parameter optimization efforts. We made significant investments in software infrastructure to enable this experiment and reduce the burden of generating future PPEs.

There were several barriers to perturbing the full set of CLM5-BGC parameters. Firstly, many parameters had not been officially identified as such. In such cases, we identified hard-coded values, established an appropriate parameter name, and extracted that parameter to the CLM parameter file for easier manipulation. In the end, we perturbed 188 parameters across a variety of CLM processes (Figure \ref{fig:params}). The second challenge involved defining a perturbation range for each parameter. We solicited expert judgment to set a minimum and maximum reasonable value for each parameter. In some cases, literature values were explicitly used (e.g. medlynslope), but the most common range was $\pm20\%$. The third challenge involved computational cost.

Running CLM-BGC with standard spin-up at 1$^\circ$ resolution, which is the most common configuration, would be prohibitively expensive, yielding only 2.3 simulations per million core-hours (for reference the CLM annual computing allocation is ~40 million core-hours). We sought a 2400-simulation ensemble, requiring a much faster model configuration. We sped up the model approximately 500x (Figure \ref{fig:sims}) by strategically reducing the number of model grid cells and by leveraging a linearized spin-up solver (see Sections \ref{sect:mcn} and \ref{sect:sg} for details). We found that most of the important model dynamics could be effectively resolved using only 400 grid cells and a 140-year spin-up sequence (Figure \ref{fig:sg}, Supp Figures \ref{supp:ilamb} and zqz spinup figure).

We found that parameter effects could be quite large, in some cases exceeding the effects of our various forcing scenarios (Figure \ref{fig:violins}). That being said, the majority of parameter perturbations had very small effect, leading to output distributions strongly concentrated around the default simulation with long tails representing a small subset of parameters. In the case of gross primary production, only xx parameters had a parameter effect of more than 2\% under the control forcing scenario (Supp Figure \ref{supp:gpp}). As such, we would expect that the dimensionality of the effective parameter space for any given process would be significantly smaller than implied by the 188 parameters.

@@

A one-at-a-time PPE cannot capture parameter interactions, and our min/max sampling protocol precludes diagnosing non-linearities. As such, this dataset will be insufficient for most calibration activities or for estimating overall parametric uncertainty. The advantage of our dataset is that we can diagnose parameter effects without the uncertainty of an approximating emulator or regression model. As such, it is easy to diagnose which parameters are most influential on a given process, which has been leveraged for parameter filtering. We have published a large set of ensemble diagnostic plots online, which serves as a valuable enhancement to our model technical documentation. Now in addition to seeing the definition of a given parameter, and the relevant equations, a model user can see the magnitude and spatial patterns of its effects (e.g. Figure \ref{fig:panel}).

>>>>

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{../figs/variance.pdf}
\caption{needs caption}
\label{fig:variance}
\end{figure}

@@

In this project, we have identified and perturbed 193 CLM parameters to create a large one-at-a-time PPE. This ensemble is useful for understanding parametric controls on CLM processes and has already been leveraged as a starting point for parameter optimization efforts. We made significant investments in software infrastructure to enable this experiment and reduce the burden of generating future PPEs.

There were several barriers to perturbing the full set of CLM5-BGC parameters. Firstly, many parameters had not been officially identified as such. In such cases, we identified hard-coded values, established an appropriate parameter name, and extracted that parameter to the CLM parameter file for easier manipulation. In the end, we perturbed 193 parameters across a variety of CLM processes (Figure \ref{fig:params}). The second challenge involved defining a perturbation range for each parameter. We solicited expert judgment to set a minimum and maximum reasonable value for each parameter. In some cases, literature values were explicitly used (e.g. medlynslope), but the most common range was $\pm20\%$.

The third challenge involved managing computational cost.  Quantifying parameter effects is a necessary prerequisite to automated calibration and uncertainty quantification. Because the parameter space of CLM is quite large and the model response potentially non-linear, large ensembles are required to adequately resolve response surfaces. With standard CLM configurations, such ensembles would far exceed our computing allocation. As such, the computational cost of inferring parameter effects is a major constraint. We were able to speed up the model approximately 500x by strategically reducing the number of model grid cells (Figure \ref{fig:sg}) and by leveraging a linearized spin-up solver (see Sections \ref{sect:mcn} and \ref{sect:sg} for details). Designing faster model configurations and more efficient sampling strategies will greatly aid in model calibration and uncertainty quantification.

For this experiment we opted for a one-at-a-time perturbation strategy, testing a minimum and maximum value for each of 193 parameters. We found that parameter effects could be quite large, in some cases exceeding the effects of our various forcing scenarios (Figure \ref{fig:violins}). That being said, the majority of parameter perturbations had very small effects, such that a majority of simulations were clustered around the default simulation. In the case of canopy evaporation, more than 90\% of the ensemble variance could be explained by the ten most influential parameters (Figure \ref{fig:variance}). As such, we would expect that the dimensionality of the effective parameter space for any given process would be significantly smaller than implied by the 193 parameters.

@@

A one-at-a-time PPE cannot capture parameter interactions, and our min/max sampling protocol precludes diagnosing non-linearities. As such, this dataset will be insufficient for most calibration activities or for estimating overall parametric uncertainty. The advantage of our dataset is that we can diagnose parameter effects without the uncertainty contributed by an  emulator or regression model. As such, it is easy to diagnose which parameters are most influential on a given process, which has already been leveraged for parameter filtering. We have published a large set of ensemble diagnostic plots online, which serves as a valuable enhancement to our model technical documentation. Now in addition to seeing the definition of a given parameter, and the relevant equations, a model user can see the magnitude and spatial patterns of its effects (e.g. Figure \ref{fig:panel}).
